On the formation of H$\alpha$ line emission around classical T Tauri star: the disc-wind-magnetosphere hybrid models

### 2006-01-29 TTau_Halpha/kurosawa.tex
Fixed the plots (Type IV-B). Now close to final version.
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@643 25464f21-6099-402c-b30a-640f1d40d38b

\usepackage{babel}
\makeatother
\begin{document}
\title[On the formation of H$\alpha$ around classical T Tauri
  stars]{On the formation of H$\alpha$ line emission around classical T Tauri stars}

@@

We adopt stellar parameters of a typical
classical T~Tauri star for the central continuum source, i.e. radius
($R_{*}$), mass ($M_{*}$), and effective temperature the photosphere
($T_{\mathrm{ph}}$) are $2\, R_{\sun}$, $0.5\, M_{\sun}$, and $4000\,\mathrm{K}$
respectively. The model atmosphere of \citet{kurucz:1979} with
$T_{\mathrm{ph}}=4000\,\mathrm{K}$ and $\log g_{*}=3.5$ (cgs) defines
the photospheric contribution to the continuum flux. The
parameters are summarised in Table~\ref{tab:std_parameters}.

@@

\begin{tabular}{lccccccccc}
\hline 
Parameters&
$R_{*}$&  
$M_{*}$&
$T_{\mathrm{ph}}$&
$R_{\mathrm{mi}}$&
$R_{\mathrm{mo}}$&
$\dot{M}_{\mathrm{acc}}$&
$\dot{M}_{\mathrm{wind}}$&
$R_{\mathrm{di}}$&
$R_{\mathrm{do}}$\tabularnewline
&
$\left[R_{\sun}\right]$&
$\left[M_{\sun}\right]$&
$\left[\mathrm{K}\right]$&
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$\left[R_{*}\right]$&
$\left[\mathrm{au}\right]$\tabularnewline
\hline
&
$2.0$&
$0.5$&
$4000$&
$2.2$&
$3.0$&
$10^{-7}$&
$10^{-8}$&
$2.2$&
$100$\tabularnewline
\hline
\end{tabular}

@@

We adopt the MA flow model of \citet{hartmann:1994},
as done  by \citet{muzerolle:2001} and by \citet*{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{fig:config_discwind} (region~2), the innermost radius ($R_{\mathrm{mi}}$)
of the magnetosphere at the equatorial plane ($z=0$) is assigned
to be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equatorial plane) are set to be $2.2\, R_{\sun}$ and $3.0\, R_{\sun}$
respectively. The geometry of the magnetic field/stream
lines is fixed for all calculations. We note that this magnetospheric geometry
is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

The magnetic field and the gas stream lines are assumed to have the
following simple form:
\begin{equation}
      r=R_{\mathrm{m}}\sin^{2}\theta
  \label{eq:dipole}
\end{equation}
\citep*[see][]{ghosh:1977} where $r$, and $\theta$ are the radial and
the polar component of a field point position vector in the accretion
stream.  $R_{\mathrm{m}}$ is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$), and its value is
restricted between $R_{\mathrm{mi}}$ and 
$R_{\mathrm{mo}}$ (Figure~\ref{fig:config_discwind}). 
Using the field geometry above and conservation of
energy, the velocity and the density of the accreting gas along the
steam line are found as in \citet{hartmann:1994}.

@@

The four basic parameters of the model are: (1)~the mass-loss rate,
(2)~the degree of the wind collimation, (3)~the velocity gradient, and
(4)~the wind temperature. The basic configuration of the disc-wind
model is shown in Figure~\ref{fig:config_discwind}. The disc wind
originates from the disc surface, but the {}``source'' point ($S$),
from which the stream lines diverges, are placed at distance $d$ above
and below the centre of the star.  The angle of the mass-loss
launching from the disc is controlled by changing the value of
$d$. The mass-loss launching occurs between $R_{\mathrm{wi}}$ and
$R_{\mathrm{wo}}$ where the former is set to be equal to the outer
radius of the magnetosphere ($R_{\mathrm{mo}}$) and the latter is set
to $1$~au as in \citet{krasnopolsky:2003}.

The local mass-loss rate per unit area ($\dot{m}$) is assumed to
be proportional to the mid-plane temperature of the disc, and is a
function of the cylindrical radius $w=\left(x^{2}+y^{2}\right)^{1/2}$, i.e.
\begin{equation}
  \dot{m}\left(w\right) \propto
       T\left(w\right)^{\alpha}\,\,.\label{eq:discwind-massloss-temp}
\end{equation}
The mid-plane temperature of the disc is assumed to be expressed
as a power-law in $w$; thus, $T\propto w^{q}$. Using this in the
relation above, one finds 
\begin{equation}
  \dot{m}\left(w\right)\propto w^{p}
  \label{eq:discwind-massloss-w}
\end{equation}
where $p=\alpha\times q$. The index of the mid-plane temperature power
law is adopted from the dust radiative transfer model of
\citet{whitney:2003a} who found the innermost part of the accretion
disc has $q=-1.15$.  In order to be consistent with the collimated
disc-wind model of \citet{krasnopolsky:2003} who used $p=-3/2$, the
value of $\alpha$ is set to 1.3. The constant of proportionality in
equation~\ref{eq:discwind-massloss-w} is found by integrating
$\dot{m}$ from $R_{\mathrm{wi}}$ to $R_{\mathrm{wo}}$, and the
normalising the value to the total mass-loss rate
$\dot{M}_{\mathrm{wind}}$.

@@

Assuming mass-flux conservation and using the velocity field defined
above, the disc wind density as a function of $w$ and $l$ can be
written as
\begin{equation}
  \rho\left(w_{i},l\right) =
  \frac{\dot{m}\left(w_{i}\right)}{v_{p}\left(w_{i},l\right)\,\left|\cos\delta\right|}\,\left\{ 
  \frac{d}{D\left(w_{i},l\right)\,\cos\delta}\right\} ^{2}
  \label{eq:discwind-density}
\end{equation} 
where $D$ and $\delta$ are the distance from the source point ($S$) to
a point along the stream line and the angle between the stream line
and the disc normal, respectively. Figure~\ref{fig:discwind_vr_rho}
shows the density and the velocity components along the mid stream
line (passing through
$w_{i}=\left(R_{\mathrm{wi}}+R_{\mathrm{wo}}\right)/2$ on the disc
plane ($z=0$) for different values of the wind acceleration parameter
$\beta$.

@@

  \caption{H$\alpha$ model profiles for wide ranges of mass accretion
    rate ($\dot{M}_\mathrm{acc}$) and temperature
    ($T_{\mathrm{max}}$). The profiles are computed using only the
    magnetospheric accretion flow (i.e. no outflow).  All the profiles
    are computed using the parameters of the reference model
    (Table~\ref{tab:std_parameters}) and inclination $i=55^{\circ}$. The
    temperature (indicated along the vertical axis) of the model increases
    from top to bottom, and the mass accretion rate (indicated by the
    values in $\mathrm{M_{\odot}\,yr^{-1}}$ along the top) increases from
    left to right. }

@@

H$\alpha$ line profiles for a range of $T_{\mathrm{max}}$ and
$\dot{M}_{\mathrm{acc}}$ are presented in  Figure~\ref{fig:atlas_acc_models}. 
The  overall dependency on $T_{\mathrm{max}}$ and $\dot{M}_{\mathrm{acc}}$ is similar to that of
\citet{muzerolle:2001}. In general, the line strength weakens
as the accretion rate and the temperature drop.  The
red-shifted absorption becomes less visible for the higher accretion rate
and higher temperature models as it becomes filled-in by the stronger Stark-broadened
line-wing emission.

We demonstrate this effect in Figure~\ref{fig:broadening}, which shows an example 
for an H$\alpha$  model with $T_{\mathrm{max}}=7500\,\mathrm{K}$
and $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$, both with and without
damping.  Although the maximum flux of the model with broadening is almost identical to that of the
model with no damping constant ($\Gamma=0$), a significant increase of
the line flux in both red and blue wings of seen. The weak red-shifted
absorption component (which is a signature of the magnetospheric accretion)
is weakened or eliminated by the flux in broadened wing.

Table~\ref{tab:ha_ew_acc_models} shows the equivalent width (EW) for
the models in Figure~\ref{fig:atlas_acc_models}.  The EWs for half the
models fall within the range of EWs ($\sim 3$~\AA\, to $\sim 160$~\AA)
measured by \citet{alencar:2000}. The EWs of models 
with the lowest mass accretion rates fall below the minimum EW
observed by  \citet{alencar:2000}, while several models would
be designated as weak-lined T~Tauri stars (WTTS) using the traditional
10\,\AA\ cut-off \citep[e.g.][]{herbig:1988}.
\citet{white:2003} empirically showed that the full width of H$\alpha$ at 10
per~cent of its peak flux (10 per~cent width) is a better
indicator for an accretion than the EW criteria. They proposed that a
T~Tauri star shows the 10 per~cent width greater than
$270\,\mathrm{km\,s^{-1}}$, the star should be a CTTS. The 10 per~cent
widths of the model H$\alpha$ profiles
(Figure~\ref{fig:atlas_acc_models}) are summarised in
Table~\ref{tab:ha_ten_percent_acc_models}.  Using the criteria of
\citet{white:2003}, the half of the models shown in the figure can be
classified as CTTS, and the other half as WTTS. However, the profiles corresponding
to the WTTS designations have an inverse P-Cygni morphology that is rarely
seen in observations (\citealt{reipurth:1996} and
section~\ref{sub:Classification-scheme-proposed}).

The dependency of the line profile on inclination angle ($i$) is
demonstrated in Figure~\ref{fig:acc_inclination_effect}. The model
has  $T_{\mathrm{max}}=7500\,\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$).  The
figure shows that the peak (normalised) flux decreases as the
inclination angle increases, as does the EW.  The line flux and the
EWs decrease mainly because the fraction of the accretion stream
blocked by the photosphere increases as the inclination increases.
The line EW in Figure~\ref{fig:acc_inclination_effect} changes from
32\,\AA\, to 21\,\AA\, as the inclination changes from $10^\circ$ to
$80^\circ$, i.e. the fractional change of 0.65.  On the other hand,
the 10~per~cent width of the line changes from 290~$\kmps$ to
470~$\kmps$; hence,
the fractional change is 1.6.  The both values are relatively
sensitive to the inclination angle.

Because of the geometry of the
magnetospheric accretion (c.f. Figure~\ref{fig:config_discwind}) and of
the presence of the gas with the highest velocity close to the stellar
surface, the highest red-shifted line-of-sight velocity is visible
only at the high inclination angles. This explains the wider
appearance of the profile with $i=80^{\circ}$ compared to the
relatively narrow line appearance of the profile with
$i=10^{\circ}$.  A similar EW dependence on
inclination angle is found for the models with different 
magnetospheric temperatures and the mass-accretion rates.

Our models show a blue-shifted line peak and
a blueward asymmetry caused by occultation of the accretion
flow by the stellar photosphere (see also \citealt{hartmann:1994};
\citealt{muzerolle:2001}). However, \citealt{alencar:2000} (see their Fig.~9)
found a substantial fraction
of the observed H$\alpha$ profiles show a red-shifted
peak. Furthermore, a recent study by
\citet{appenzeller:2005} demonstrated  that the equivalent width of H$\alpha$
from CTTS {\em increases} as the inclination angle increases. 
Clearly a magnetospheric accretion only model cannot explain the
full properties of the H$\alpha$ profiles in CTTS.

@@

\caption{Summary of H$\alpha$ equivalent widths (\AA) from the
magnetospheric accretion flow models shown in
Figure~\ref{fig:atlas_acc_models}. Note that the equivalent widths are
positive when the lines are in emission.}

@@

\caption{Summary of 10 per~cent widths ($\mathrm{km\,s^{-1}}$) from the
magnetospheric accretion flow models shown in
Figure~\ref{fig:atlas_acc_models}.}

@@

Initially we examine the characteristics of the H$\alpha$ profile as a
function of the wind acceleration parameter $\beta$ and the isothermal
wind temperature $T_{\mathrm{wind}}$.  The mass-loss rate is kept
constant at
$\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\,yr^{-1}}$.
Figure~\ref{fig:atlas_discwind_only} shows the model profiles computed
for the wind temperatures between $7000\,\mathrm{K}$ and
$9000\,\mathrm{K}$, and those with $\beta$ between $0.2$ and $2.0$,
computed at inclination $i=55^{\circ}$.
The morphology of the profile exhibited by the model changes from
FU-Ori type (c.f.~\citealt{reipurth:1996}) for the models with small
$\beta$ to the stronger emission types, which are more commonly seen in
the observed H$\alpha$ (c.f.~\citealt{reipurth:1996}), as the value of
$\beta$ increases.  
Classical P~Cygni profiles are prominent
in the models with lower values of $T_{\mathrm{wind}}$ and $\beta$
(a colder and faster accelerating wind). As the value of $\beta$
increases, the position of the blue-shifted absorption component moves
toward the line centre. The blue edge of the absorption component in
the fastest accelerating wind 
($\beta=0.2$) model is almost at the terminal velocity ($v_{\infty}\sim
160\,\kmps$) of the disc wind (c.f.~Figure~\ref{fig:discwind_vr_rho}), that
for the slowest wind acceleration model 
($\beta=4.0$) is located close the line centre ($\sim 20\,\kmps$).

The intensity of the emission component increases as
$\beta$ increases for a fixed $T_{\mathrm{wind}}$. This is mainly
because the density near the photosphere increases as the value of
$\beta$ increases (Figure~\ref{fig:discwind_vr_rho}); hence, the
emissivity of increases. Most of the profiles show a red-ward asymmetry, mainly
due to the presence of the classical P-Cygni absorption. However we find
that the disc-wind models with the fastest acceleration ($\beta=0.2$ and $0.5$)
and viewed at low inclinations, show a strong blue asymmetry due to
occultation of the receding half of the outflow by the optically thick disc.

@@

  \caption{The dependency of the H$\alpha$ profile on the source
    displacement parameter $d$. The line profiles are computed with 
    the wind parameter $T_{\mathrm{wind}} = 9000\Kelvin$ and $\beta  =
    0.5$ for $i=55^{\circ}$.  The profiles in the figure use $d = 
    11~R_{*}$ (dashed),   $18~R_{*}$ (dash-dot) and $22~R_{*}$
    (dash-dot). All the other parameters are same as in
    Figure~\ref{fig:atlas_discwind_only}.  As the wind becomes more
    collimated (larger $d$), the wind launching angle becomes smaller
    and the wind density is enhanced towards the symmetry axis
    ($z$); hence, the wind emission increases.  The absorption
    component is relatively insensitive to $d$ since the change in the
    column density toward the observer is relatively small. }

@@

@@

The effect of varying the wind mass-loss rate is shown in
Figure~\ref{fig:discwind_mdot_effect}. With all other parameters fixed,
$\dot{M}_{\mathrm{wind}}$ acts as an scaling factor for the wind
density (equation~\ref{eq:discwind-density}). The important process
that populates the $n=3$ level is recombination, and therefore the
line emissivity varies as the square of the density (assuming that the
ionisation fraction remains constant); hence the line is very sensitive to 
changes in $\dot{M}_{\mathrm{wind}}$.  We indeed find that  the line flux of
the models is approximately proportional to the square
of the mass-loss rate.

@@

  \caption{The disc-wind-magnetosphere hybrid model. These models
  (solid lines) have the  same parameters as
    Figure~\ref{fig:atlas_discwind_only}, but  also
    include the magnetospheric accretion flow
    ($T_{\mathrm{\max}}=7500~\mathrm{K}$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$). While
    the emission from the magnetosphere dominates for the models with
    smaller $\beta$ and $T_{\mathrm{wind}}$, the emission from the wind dominates the
    profiles for models with larger $\beta$ and $T_{\mathrm{wind}}$.  The contribution
    of the magnetosphere can be judged from the profiles computed with
    the magnetosphere only (dotted lines).}

@@

xxx add a sample halpha image from poster xxxx

@@

  \caption{An H$\alpha$ image of the disc-wind hybrid model viewed at an inclination
  of 80$^\circ$. The false-colour scale shows logarithmic surface brightness (arbitrary
  units) and the axes are linear dimensions in au.}

@@

\section{H$\alpha$ atlas of the disc-wind-magnetosphere hybrid model}

>>>>

\usepackage{babel}
\makeatother
\begin{document}
\title[On the formation of H$\alpha$ around classical T Tauri
  stars]{On the formation of H$\alpha$ line emission around classical
  T Tauri star: the disc-wind-magnetosphere hybrid models}

@@

We adopt stellar parameters of a typical
classical T~Tauri star for the central continuum source, i.e. radius
($R_{*}$), mass ($M_{*}$), and effective temperature the photosphere
($T_{\mathrm{ph}}$) are $2\, \mathrm{R_{\sun}}$, $0.5\,
\mathrm{M_{\sun}}$, and $4000\,\mathrm{K}$ 
respectively. The model atmosphere of \citet{kurucz:1979} with
$T_{\mathrm{ph}}=4000\,\mathrm{K}$ and $\log g_{*}=3.5$ (cgs) defines
the photospheric contribution to the continuum flux. The
parameters are summarised in Table~\ref{tab:std_parameters}.

@@

\begin{tabular}{lccccccccc}
\hline 
Parameters&
$R_{*}$&  
$M_{*}$&
$T_{\mathrm{ph}}$&
$R_{\mathrm{mi}}$&
$R_{\mathrm{mo}}$&
$\dot{M}_{\mathrm{acc}}$&
$\dot{M}_{\mathrm{wind}}$&
$R_{\mathrm{di}}$&
$R_{\mathrm{do}}$\tabularnewline
&
$\left[\mathrm{R_{\sun}}\right]$&
$\left[\mathrm{M_{\sun}}\right]$&
$\left[\mathrm{K}\right]$&
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[\mathrm{M_{\sun}}\,\mathrm{yr^{-1}}\right]$&
$\left[\mathrm{M_{\sun}}\,\mathrm{yr^{-1}}\right]$&
$\left[R_{*}\right]$&
$\left[\mathrm{au}\right]$\tabularnewline
\hline
&
$2.0$&
$0.5$&
$4000$&
$2.2$&
$3.0$&
$10^{-7}$&
$10^{-8}$&
$2.2$&
$100$\tabularnewline
\hline
\end{tabular}

@@

We adopt the MA flow model of \citet{hartmann:1994},
as done  by \citet{muzerolle:2001} and by \citet*{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{fig:config_discwind} (region~2), the innermost radius ($R_{\mathrm{mi}}$)
of the magnetosphere at the equatorial plane ($z=0$) is assigned
to be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equatorial plane) are set to be $2.2\, \mathrm{R_{\sun}}$ and $3.0\, \mathrm{R_{\sun}}$
respectively. The geometry of the magnetic field/stream
lines is fixed for all calculations. We note that this magnetospheric geometry
is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

The magnetic field and the gas stream lines are assumed to have the
following simple form:
\begin{equation}
      r=R_{\mathrm{m}}\sin^{2}\theta
  \label{eq:dipole}
\end{equation}
\citep*[see][]{ghosh:1977} where $r$, and $\theta$ are the radial and
the polar component of a field point position vector in the accretion
stream.  $R_{\mathrm{m}}$ is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$), and its value is
restricted between $R_{\mathrm{mi}}$ and 
$R_{\mathrm{mo}}$ (Fig.~\ref{fig:config_discwind}). 
Using the field geometry above and conservation of
energy, the velocity and the density of the accreting gas along the
steam line are found as in \citet{hartmann:1994}.

@@

The four basic parameters of the model are: (1)~the mass-loss rate,
(2)~the degree of the wind collimation, (3)~the velocity gradient, and
(4)~the wind temperature. The basic configuration of the disc-wind
model is shown in Fig.~\ref{fig:config_discwind}. The disc wind
originates from the disc surface, but the {}``source'' point ($S$),
from which the stream lines diverges, are placed at distance $d$ above
and below the centre of the star.  The angle of the mass-loss
launching from the disc is controlled by changing the value of
$d$. The mass-loss launching occurs between $R_{\mathrm{wi}}$ and
$R_{\mathrm{wo}}$ where the former is set to be equal to the outer
radius of the magnetosphere ($R_{\mathrm{mo}}$) and the latter is set
to $1$~au as in \citet{krasnopolsky:2003}.

The local mass-loss rate per unit area ($\dot{m}$) is assumed to
be proportional to the mid-plane temperature of the disc, and is a
function of the cylindrical radius $w=\left(x^{2}+y^{2}\right)^{1/2}$, i.e.
\begin{equation}
  \dot{m}\left(w\right) \propto
       \left[ T\left(w\right) \right]  ^{\alpha}\,\,.\label{eq:discwind-massloss-temp}
\end{equation}
The mid-plane temperature of the disc is assumed to be expressed
as a power-law in $w$; thus, $T\propto w^{q}$. Using this in the
relation above, one finds 
\begin{equation}
  \dot{m}\left(w\right)\propto w^{p}
  \label{eq:discwind-massloss-w}
\end{equation}
where $p=\alpha\times q$. The index of the mid-plane temperature power
law is adopted from the dust radiative transfer model of
\citet{whitney:2003a} who found the innermost part of the accretion
disc has $q=-1.15$.  In order to be consistent with the collimated
disc-wind model of \citet{krasnopolsky:2003} who used $p=-3/2$, the
value of $\alpha$ is set to 1.3. The constant of proportionality in
equation~\ref{eq:discwind-massloss-w} is found by integrating
$\dot{m}$ from $R_{\mathrm{wi}}$ to $R_{\mathrm{wo}}$, and the
normalising the value to the total mass-loss rate
$\dot{M}_{\mathrm{wind}}$.

@@

Assuming mass-flux conservation and using the velocity field defined
above, the disc wind density as a function of $w$ and $l$ can be
written as
\begin{equation}
  \rho\left(w_{i},l\right) =
  \frac{\dot{m}\left(w_{i}\right)}{v_{p}\left(w_{i},l\right)\,\left|\cos\delta\right|}\,\left\{ 
  \frac{d}{D\left(w_{i},l\right)\,\cos\delta}\right\} ^{2}
  \label{eq:discwind-density}
\end{equation} 
where $D$ and $\delta$ are the distance from the source point ($S$) to
a point along the stream line and the angle between the stream line
and the disc normal, respectively. Fig.~\ref{fig:discwind_vr_rho}
shows the density and the velocity components along the mid stream
line (passing through
$w_{i}=\left(R_{\mathrm{wi}}+R_{\mathrm{wo}}\right)/2$ on the disc
plane ($z=0$) for different values of the wind acceleration parameter
$\beta$.

@@

  \caption{H$\alpha$ model profiles for wide ranges of mass accretion
    rate ($\dot{M}_\mathrm{acc}$) and temperature
    ($T_{\mathrm{max}}$). The profiles are computed using only the
    magnetospheric accretion flow (i.e. no outflow).  All the profiles
    are computed using the parameters of the reference model
    (Table~\ref{tab:std_parameters}) and inclination $i=55^{\circ}$. The
    temperature (indicated along the vertical axis) of the model increases
    from top to bottom, and the mass accretion rate (indicated by the
    values in $\mathrm{M_{\sun}\,yr^{-1}}$ along the top) increases from
    left to right. }

@@

H$\alpha$ line profiles for a range of $T_{\mathrm{max}}$ and
$\dot{M}_{\mathrm{acc}}$ are presented in  Fig.~\ref{fig:atlas_acc_models}. 
The  overall dependency on $T_{\mathrm{max}}$ and $\dot{M}_{\mathrm{acc}}$ is similar to that of
\citet{muzerolle:2001}. In general, the line strength weakens
as the accretion rate and the temperature drop.  The
red-shifted absorption becomes less visible for the higher accretion rate
and higher temperature models as it becomes filled-in by the stronger Stark-broadened
line-wing emission.

We demonstrate this effect in Fig.~\ref{fig:broadening}, which shows an example 
for an H$\alpha$  model with $T_{\mathrm{max}}=7500\,\mathrm{K}$
and $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$, both with and without
damping.  Although the maximum flux of the model with broadening is almost identical to that of the
model with no damping constant ($\Gamma=0$), a significant increase of
the line flux in both red and blue wings of seen. The weak red-shifted
absorption component (which is a signature of the magnetospheric accretion)
is weakened or eliminated by the flux in broadened wing.

Table~\ref{tab:ha_ew_acc_models} shows the equivalent width (EW) for
the models in Fig.~\ref{fig:atlas_acc_models}.  The EWs for half the
models fall within the range of EWs ($\sim 3$~\AA\, to $\sim 160$~\AA)
measured by \citet{alencar:2000}. The EWs of models 
with the lowest mass accretion rates fall below the minimum EW
observed by  \citet{alencar:2000}, while several models would
be designated as weak-lined T~Tauri stars (WTTS) using the traditional
10\,\AA\ cut-off \citep[e.g.][]{herbig:1988}.
\citet{white:2003} empirically showed that the full width of H$\alpha$ at 10
per~cent of its peak flux (10 per~cent width) is a better
indicator for an accretion than the EW criteria. They proposed that a
T~Tauri star shows the 10 per~cent width greater than
$270\,\mathrm{km\,s^{-1}}$, the star should be a CTTS. The 10 per~cent
widths of the model H$\alpha$ profiles
(Fig.~\ref{fig:atlas_acc_models}) are summarised in
Table~\ref{tab:ha_ten_percent_acc_models}.  Using the criteria of
\citet{white:2003}, the half of the models shown in the figure can be
classified as CTTS, and the other half as WTTS. However, the profiles corresponding
to the WTTS designations have an inverse P-Cygni morphology that is rarely
seen in observations (\citealt{reipurth:1996} and
section~\ref{sub:Classification-scheme-proposed}).

The dependency of the line profile on inclination angle ($i$) is
demonstrated in Fig.~\ref{fig:acc_inclination_effect}. The model
has  $T_{\mathrm{max}}=7500\,\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$).  The
figure shows that the peak (normalised) flux decreases as the
inclination angle increases, as does the EW.  The line flux and the
EWs decrease mainly because the fraction of the accretion stream
blocked by the photosphere increases as the inclination increases.
The line EW in Fig.~\ref{fig:acc_inclination_effect} changes from
32\,\AA\, to 21\,\AA\, as the inclination changes from $10^\circ$ to
$80^\circ$, i.e. the fractional change of 0.65.  On the other hand,
the 10~per~cent width of the line changes from 290~$\kmps$ to
470~$\kmps$; hence,
the fractional change is 1.6.  The both values are relatively
sensitive to the inclination angle.

Because of the geometry of the
magnetospheric accretion (c.f. Fig.~\ref{fig:config_discwind}) and of
the presence of the gas with the highest velocity close to the stellar
surface, the highest red-shifted line-of-sight velocity is visible
only at the high inclination angles. This explains the wider
appearance of the profile with $i=80^{\circ}$ compared to the
relatively narrow line appearance of the profile with
$i=10^{\circ}$.  A similar EW dependence on
inclination angle is found for the models with different 
magnetospheric temperatures and the mass-accretion rates.

Our models show a blue-shifted line peak and
a blueward asymmetry caused by occultation of the accretion
flow by the stellar photosphere (see also \citealt{hartmann:1994};
\citealt{muzerolle:2001}). However, \citealt{alencar:2000} (see their fig.~9)
found a substantial fraction
of the observed H$\alpha$ profiles show a red-shifted
peak. Furthermore, a recent study by
\citet{appenzeller:2005} demonstrated  that the equivalent width of H$\alpha$
from CTTS {\em increases} as the inclination angle increases. 
Clearly a magnetospheric accretion only model cannot explain the
full properties of the H$\alpha$ profiles in CTTS.

@@

\caption{Summary of H$\alpha$ equivalent widths (\AA) from the
magnetospheric accretion flow models shown in
Fig.~\ref{fig:atlas_acc_models}. Note that the equivalent widths are
positive when the lines are in emission.}

@@

\caption{Summary of 10 per~cent widths ($\mathrm{km\,s^{-1}}$) from the
magnetospheric accretion flow models shown in
Fig.~\ref{fig:atlas_acc_models}.}

@@

Initially we examine the characteristics of the H$\alpha$ profile as a
function of the wind acceleration parameter $\beta$ and the isothermal
wind temperature $T_{\mathrm{wind}}$.  The mass-loss rate is kept
constant at
$\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\,yr^{-1}}$.
Fig.~\ref{fig:atlas_discwind_only} shows the model profiles computed
for the wind temperatures between $7000\,\mathrm{K}$ and
$9000\,\mathrm{K}$, and those with $\beta$ between $0.2$ and $2.0$,
computed at inclination $i=55^{\circ}$.
The morphology of the profile exhibited by the model changes from
FU-Ori type (c.f.~\citealt{reipurth:1996}) for the models with small
$\beta$ to the stronger emission types, which are more commonly seen in
the observed H$\alpha$ (c.f.~\citealt{reipurth:1996}), as the value of
$\beta$ increases.  
Classical P~Cygni profiles are prominent
in the models with lower values of $T_{\mathrm{wind}}$ and $\beta$
(a colder and faster accelerating wind). As the value of $\beta$
increases, the position of the blue-shifted absorption component moves
toward the line centre. The blue edge of the absorption component in
the fastest accelerating wind 
($\beta=0.2$) model is almost at the terminal velocity ($v_{\infty}\sim
160\,\kmps$) of the disc wind (c.f.~Fig.~\ref{fig:discwind_vr_rho}), that
for the slowest wind acceleration model 
($\beta=4.0$) is located close the line centre ($\sim 20\,\kmps$).

The intensity of the emission component increases as
$\beta$ increases for a fixed $T_{\mathrm{wind}}$. This is mainly
because the density near the photosphere increases as the value of
$\beta$ increases (Fig.~\ref{fig:discwind_vr_rho}); hence, the
emissivity of increases. Most of the profiles show a red-ward asymmetry, mainly
due to the presence of the classical P-Cygni absorption. However we find
that the disc-wind models with the fastest acceleration ($\beta=0.2$ and $0.5$)
and viewed at low inclinations, show a strong blue asymmetry due to
occultation of the receding half of the outflow by the optically thick disc.

@@

  \caption{The dependency of the H$\alpha$ profile on the source
    displacement parameter $d$. The line profiles are computed with 
    the wind parameter $T_{\mathrm{wind}} = 9000\Kelvin$ and $\beta  =
    0.5$ for $i=55^{\circ}$.  The profiles in the figure use $d = 
   2~R_{*}$ (dashed),   $7~R_{*}$ (dash-dot) and $88~R_{*}$
    (dash-dot). All the other parameters are same as in
    Fig.~\ref{fig:atlas_discwind_only}.  As the wind becomes more
    collimated (larger $d$), the wind emission becomes stronger.  
    The P-Cygni absorption disappears for the model with a small
    collimation (e.g.~$d = 2~R_{*}$ case) since the opening angel of
    the disc wind is larger than the inclination angle. }

@@

We now examine the sensitivity of the model profiles on the wind
collimation or the source displacement parameter $d$. As we can see
from Fig.~\ref{fig:config_discwind}, the disc wind
geometry/collimation is determined by $d$ and the wind launching radii
($R_{\mathrm{wi}}$ and $R_{\mathrm{wo}}$). Here, we fix the values of
$R_{\mathrm{wi}}$ and $R_{\mathrm{wo}}$ as in
Table~\ref{tab:std_discwind_parameters}, and inspect the model
dependency on $d$ alone.  The minimum wind opening angle (measured
from the symmetry axis) is given by
$\theta_{\mathrm{open}}=\arctan\left( d/R_{\mathrm{wi}} \right)$.
Fig.~\ref{fig:discwind_d_effect} shows the profiles computed for $d
= 2~R_{*}$, $7~R_{*}$ and $88~R_{*}$ for the inclination angle
$i=55^{\circ}$.  As the wind becomes more collimated (larger $d$), the
wind launching angle becomes smaller and the wind density is enhanced
towards the symmetry axis ($z$); hence, the wind emission increases.
When the collimation of the wind is small (as in the $d = 2~R_{*}$
case), the P-Cygni absorption is absent from the profile because the
opening angel of the disc wind is greater than the inclination angle.
For the disc wind models presented in this paper, we use $d=44~R_{*}$
to have a moderate amount of wind absorption at a moderate inclination
angle ($i\sim60^\circ$).

@@

The effect of varying the wind mass-loss rate is shown in
Fig.~\ref{fig:discwind_mdot_effect}. With all other parameters fixed,
$\dot{M}_{\mathrm{wind}}$ acts as an scaling factor for the wind
density (equation~\ref{eq:discwind-density}). The important process
that populates the $n=3$ level is recombination, and therefore the
line emissivity varies as the square of the density (assuming that the
ionisation fraction remains constant); hence the line is very sensitive to 
changes in $\dot{M}_{\mathrm{wind}}$.  We indeed find that  the line flux of
the models is approximately proportional to the square
of the mass-loss rate.

@@

  \caption{The disc-wind-magnetosphere hybrid model. These models
  (solid lines) have the  same parameters as
    Fig.~\ref{fig:atlas_discwind_only}, but  also
    include the magnetospheric accretion flow
    ($T_{\mathrm{\max}}=7500~\mathrm{K}$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$). While
    the emission from the magnetosphere dominates for the models with
    smaller $\beta$ and $T_{\mathrm{wind}}$, the emission from the wind dominates the
    profiles for models with larger $\beta$ and $T_{\mathrm{wind}}$.  The contribution
    of the magnetosphere can be judged from the profiles computed with
    the magnetosphere only (dotted lines).}

@@

@@

  \caption{An H$\alpha$ image of the disc-wind hybrid model viewed at an inclination
  of 80$^\circ$. The false-colour scale shows logarithmic (base 10)
  surface brightness (arbitrary units) and the axes are linear
  dimensions in au. The parameter used for the magnetosphere are as
  the reference model in Table~\ref{tab:std_parameters}. The wind
  temperature of $T_{\mathrm{max}}=7000~\Kelvin$ and the wind
  acceleration parameter $\beta=1.0$ are adopted here.  }

@@

\section{H$\alpha$ atlas of the disc-wind-magnetosphere hybrid model}
\label{sec:appendix}
